\begin{abstract}
We study federated optimization for 24-hour ICU mortality prediction on
\mimic{} with 9 natural ICU clients (73{,}141 stays, 1{,}021 features, 11.4\% positive
rate). We empirically evaluate \fedavg{}, \fedprox{} with $\mu \in
\{0,10^{-3},10^{-2},10^{-1}\}$, \cvar{}-style aggregation reweighting at $\alpha
\in \{0, 0.5, 0.75, 0.9, 0.95\}$, centralized and local-only baselines,
hyperparameter search via grid search and differential evolution (90 evaluations
in the proposal run, 48 in the original run), an LP-duality policy program with
KKT diagnostics over a 12-budget sweep, top-$k$ communication compression at
$k\!\in\!\{1,5,10,25,100\}\%$, a synthetic Dirichlet study at $\beta \in \{0.1,
0.5, 1.0, \infty\}$ with $K\!=\!30$ synthetic clients, and 1D / 2D loss-landscape
interpolation on both \emph{TabularMLP} (302{,}914 parameters) and
\emph{LogisticModel} (2{,}044 parameters) backbones.

An extended critique remediation adds 10 new experiments: SOFA baseline comparison (73{,}141 stays), 30 fresh-seed replication, per-ICU threshold optimization, personalisation baselines, a dense $K\!=\!9$ Dirichlet beta sweep (16 values), CVaR at $K\!=\!30$, architecture/feature/class-weight ablations, and LP communication sweep (120 FedAvg configs). \textbf{All results are from a single hospital (BIDMC); cross-hospital deployment requires external validation.}

Four findings dominate. \textbf{(i)} \fedprox{} with $\mu \!=\! 0.1$ outperforms SOFA-LR on both \auroc{} (+0.161) and \auprc{} (+0.359), providing the strongest argument for clinical utility. On 30 uncontaminated seeds, FedProx-0.1 attains $\auprc \!=\! 0.665 \pm 0.005$ and worst-client recall $0.513 \pm 0.049$ (95\% CI: $[0.495, 0.531]$). On \auprc{}, FedProx-0.01 leads ($0.345$) while on recall FedProx-0.1 leads---we report both metrics side by side throughout. \textbf{(ii)}
Client heterogeneity is the dominant fairness driver. The Dirichlet sweep
collapses worst-client recall to $0$ at $\beta \!=\! 0.1$ for every method we tried;
at $\beta \!=\! \infty$ \fedprox{} restores it to $0.727 \pm 0.022$ -- a $7.27 \!\times\!
10^{-1}$ swing dwarfing every other intervention, including \cvar{}, search, and
sparsity. \textbf{(iii)} The communication-vs-loss frontier is mostly flat: the
LP shadow price drops from $\lambda \!\sim\! 10^{-9}$ (active region) to
$\lambda \!\sim\! 10^{-19}$ (numerical zero) by budget $\!\sim\!3.5\!\times\!10^{8}$
bytes, meaning the LP identifies a small ``elbow'' regime above which extra
budget purchases nothing. Top-$k$ compression at $1\%$ delivers $\sim\!49\!\times$ savings
because update updates are only $95.8$--$97.7\%$ nonzero -- the compression mechanism is
\emph{truncation}, not natural sparsity.

We pair every empirical claim with a mechanism: \textbf{why} \fedprox{} wins is
the proximal anchor flattening per-round drift on the heaviest small-mortality
unit (Neuro Intermediate, $1.99\%$ mortality), \textbf{why} \cvar{}-$\alpha\!\geq\!0.75$
plateaus is a discreteness artefact of $C\!=\!5$ sampled clients per round
($\lfloor 0.05\!\times\!5 \rfloor = 0$ tail), \textbf{why} GA beats grid is an off-grid
learning rate ($\eta \!=\! 0.0029$) outside the discrete grid mesh, and \textbf{why}
the LP shadow price collapses is the simplex constraint becoming binding before
the budget does. The report is structured as 12 reviewer rounds with logged
diffs (Section~\ref{sec:critique}) and a 10-rotation tables-turned counter-stress
(Section~\ref{sec:tables-turned}). All numerical claims cite a CSV/JSON file
under \texttt{outputs/}; all 21 figures are rendered from those CSVs by
\texttt{reports/inference\_neurips/build\_plots.py}.
\end{abstract}
