\section{Background and notation}
\label{sec:background}

\paragraph{Federated learning, briefly.} Let $K$ clients each hold a private
training set $\mathcal{D}_k$ with empirical risk $F_k(w) = \frac{1}{|\mathcal{D}_k|}
\sum_{(x,y)\in\mathcal{D}_k} \ell(w; x, y)$. The global objective most federated
algorithms target is the size-weighted population risk
\[
F(w) \;=\; \sum_{k=1}^{K} \frac{|\mathcal{D}_k|}{N} \, F_k(w),
\qquad N \;=\; \sum_k |\mathcal{D}_k|.
\]
\fedavg{}~[McMahan et al., 2017] samples a subset $S_t \subseteq \{1,\dots,K\}$ of
size $C$ each round $t$, runs $E$ epochs of local SGD on each selected client,
and aggregates by size weights:
\[
w_{t+1} \;=\; \sum_{k \in S_t} \frac{|\mathcal{D}_k|}{\sum_{j\in S_t} |\mathcal{D}_j|}
            \, w_{t,k}^{(E)}.
\]
For our \mimic{} setup, $K = 9$, $C = 5$ per round
(\texttt{run\_metadata.json:clients\_per\_round}), $E = 1$, batch size $256$,
optimizer Adam with default $\beta_1 = 0.9, \beta_2 = 0.999$. The early-stopping
rule is patience $25$ rounds, $\Delta_{\min} = 5\!\times\!10^{-4}$ on the mean
held-out loss, with best-state checkpointing
(\texttt{flopt/fedavg.py:18-72}).

\paragraph{The four trade-off axes.} Throughout we will treat federated learning
on \mimic{} as a four-axis trade-off. None of the four axes are dominated by
a single method:
\[
\underbrace{\text{Accuracy}}_{(\text{axis 1})}
\;\leftrightarrow\;
\underbrace{\auprc{}}_{(\text{axis 2})}
\;\leftrightarrow\;
\underbrace{\worstrec{} \text{ (worst-client recall)}}_{(\text{axis 3})}
\;\leftrightarrow\;
\underbrace{\text{Comm}}_{(\text{axis 4})}.
\]
Section~\ref{sec:results} shows the leaderboards for axes 1--4 are essentially
disjoint. The bulk of this report's inference work is explaining \emph{why} no
method dominates.

\paragraph{The two backbones.} We use exactly two model architectures, kept
small on purpose because the entire experimental matrix has to fit in 48~GB
unified memory on Apple Silicon (\texttt{monitoring/resource\_summary.csv}):
\begin{itemize}[leftmargin=*,topsep=2pt]
    \item \textbf{TabularMLP} (\texttt{flopt/models.py:TabularMLP}): $1021 \!\to\! 256
        \!\to\! 128 \!\to\! 64 \!\to\! 2$, ReLU, dropout $0.1$, weighted CE.
        Parameter count
        $1021\!\times\!256 + 256 + 256\!\times\!128 + 128 + 128\!\times\!64 + 64 + 64\!\times\!2 + 2 = 302{,}914$
        (matches \texttt{run\_metadata.json:model\_parameters}).
    \item \textbf{LogisticModel} (\texttt{flopt/models.py:LogisticModel}): a single
        $1021 \!\to\! 2$ linear layer, $1021\!\times\!2 + 2 = 2{,}044$
        parameters. $148\!\times$ smaller in upload bytes per round.
\end{itemize}
Logistic is convex; MLP is non-convex. The design rationale is that anything
true on both backbones survives non-convexity; anything true on MLP only is
suspect.

\paragraph{Headline metrics.} For each method/seed we record:
\begin{itemize}[leftmargin=*,nosep]
    \item Accuracy $= (TP + TN) / (TP + TN + FP + FN)$
    \item \auroc{} (rank-based, threshold-free, robust to imbalance)
    \item \auprc{} (precision-recall area, the \emph{primary} metric for $11.4\%$
        positive rate)
    \item Sensitivity (recall on the \emph{death} class -- clinically the
        north-star)
    \item Specificity (recall on the \emph{survival} class)
    \item Worst-client recall: $\worstrec{} = \min_{k\in[K]} \text{Sens}_k$, the
        smallest death-recall across the 9 ICUs. The clinical fairness metric.
    \item Total upload-plus-download bytes until early-stop: $4 \times P \times C
        \times 2 \times R$ where $P$ is parameter count and $R$ is the
        \texttt{stopped\_round}.
\end{itemize}
Why this list and not others: in a $11.4\%$-positive cohort, mean accuracy is
low-information (the trivial baseline is at $88.6\%$); \auroc{} hides class
imbalance; only \auprc{} reflects how well rare-class probability is ranked.
Worst-client recall is the deployment metric -- if a federated model is rolled
to all $9$ ICUs and the smallest one (Neuro Intermediate, $10$ deaths in the
test set) catches $0/10$, that ICU's clinicians correctly distrust the model
even if global \auroc{} is $0.91$.

\paragraph{The federated objective is not the centralized objective.}
A subtle but central point: under heterogeneous data, $\arg\min_w F(w)$ is
\emph{not} the empirical risk minimum on the pooled dataset, because each
$F_k$ has a different optimum and the federated aggregation is essentially a
$|\mathcal{D}_k|$-weighted convex combination of per-client gradients. \fedprox{}
modifies this by adding a proximal anchor $\frac{\mu}{2} \|w - w_t\|_2^2$ to
each $F_k$ during the local update, which biases each client toward the global
$w_t$ \emph{before} aggregation -- a key step in why \fedprox{} on this dataset
both improves \auprc{} and lowers communication
(Section~\ref{sec:mechanism-fedprox}).
