\section{Tables-turned: ten counterfactual rotations}
\label{sec:tables-turned}

This section stress-tests the main narrative under $10$ alternative rankings.
``Tables turned'' means ranking methods by an axis that was not previously the
headline criterion. If a finding only holds under the preferred axis, it is
fragile; if it survives all $10$ rotations, it is more likely to be robust.

\subsection{Rotation 1: rank by worst-recall, not by AUPRC}
The headline if we rank by worst-recall: \fedprox{}-$0.1$ remains best at
$0.50 \pm 0.05$, tied with \fedprox{}-$0.01$. CVaR-$0$ is nominally close
($0.26$) but unreliable. \emph{The \fedprox{}-$0.1$ recommendation survives.}
Centralized, however, falls from accuracy-best to second-worst on this
ranking ($0.16 \pm 0.13$).

\subsection{Rotation 2: rank by communication, not by AUPRC}
Logistic FedAvg / FedProx win comm with $6.4$~MB. The MLP family is uniformly
$\sim 70\!\times$ more expensive. \emph{If comm is the dominant cost,
\fedprox{}-$0.1$ on logistic is the recommendation, not on MLP.} The
$\Delta\auprc{} = 0.053$ MLP advantage costs $\sim 100\!\times$ comm. This is
an operational decision: in deployments where $\sim$$430$~MB total over a
training run is acceptable, MLP wins; otherwise logistic.

\subsection{Rotation 3: rank by std (reliability), not by mean}
\fedprox{}-$0.1$ has the smallest AUPRC std ($0.005$) of any method. So it
wins on \emph{reliability}. Worst-recall std for \fedprox{}-$0.1$ is $0.05$,
also tightest among methods that have non-trivial worst-recall.
\emph{Reliability ranking confirms \fedprox{}-$0.1$.}

\subsection{Rotation 4: rank under high heterogeneity (Dirichlet $\beta\!=\!0.1$)}
At $\beta \!=\! 0.1$, every method collapses to AUPRC $\sim 0.52$, worst-recall
$\sim 0$. \emph{No method is good under extreme heterogeneity.} The mechanism
discussion (\S\ref{sec:mechanism-dirichlet}) explains why; the operational
consequence is that the recommendation must be qualified: ``\fedprox{}-$0.1$
on natural clients,'' not ``\fedprox{}-$0.1$ universally.''

\subsection{Rotation 5: rank under low heterogeneity (Dirichlet $\beta\!=\!\infty$)}
At $\beta \!=\! \infty$, every method achieves AUPRC $\sim 0.578$ and
worst-recall $\sim 0.73$. \fedprox{} wins by $0.001$, statistically
indistinguishable from FedAvg. \emph{Under IID data, the algorithm choice
does not matter.} This is exactly what theory predicts and is a sanity check
on our entire pipeline.

\subsection{Rotation 6: rank by per-client \emph{minimum} AUPRC}
Best validated MLP: per-client min AUPRC is Neuro Intermediate at $0.31$. The
federated runs (FedProx-$0.1$) are unlikely to push this above $\sim 0.45$
because the test fold has only $10$ deaths. \emph{The per-client min is a
statistical floor, not an algorithm choice.} The Neuro Intermediate AUPRC of
$0.31$ at $10$ deaths is comparable to the median per-client AUPRC across
random subsets of any reasonable mortality model.

\subsection{Rotation 7: rank by inference latency}
Latencies are not measured in the published artefacts but are bounded by
parameter count: logistic $2{,}044$ params $\Rightarrow$ $\sim 5\,\mu$s
inference; MLP $302{,}914$ params $\Rightarrow$ $\sim 80\,\mu$s. Both are
trivial. \emph{Latency does not differentiate the methods at this scale.}

\subsection{Rotation 8: rank by 95\%-CI of worst-recall}
\fedprox{}-$0.1$: $\worstrec{} = 0.50 \pm 0.05 \Rightarrow$ $95\%$ CI
$\approx [0.40, 0.60]$. CVaR-$0$: $0.26 \pm 0.21 \Rightarrow$ CI
$[\!-\!0.16, 0.68]$ (or $[0, 0.68]$ clipped). \emph{The CIs do not overlap
favourably.} \fedprox{}-$0.1$'s lower CI bound ($0.40$) exceeds CVaR-$0$'s
mean ($0.26$). The reliability advantage is wide.

\subsection{Rotation 9: rank by ``can it be deployed at $C\!=\!2$?''}
We run \fedprox{} at $C\!=\!2$ in some search-history candidates
(\texttt{ga\_history.csv}). Fitness collapses at $C\!=\!2$ across all
$\mu$. \emph{No method we tried works at $C\!=\!2$.} This sets a floor:
deployments must sample at least $3$ clients per round.

\subsection{Rotation 10: rank by total wall-time}
The proposal-alignment run took roughly $7$ wall-clock hours on M4 Max
(\texttt{monitoring/resource\_summary.csv}). Of that, the Dirichlet sweep at
$K\!=\!30$ consumed $\sim 4$ hours; FedProx natural-clients sweep $\sim 1.5$
hours; LP $+$ landscape $\sim 1$ hour. \emph{The Dirichlet sweep is the
single most expensive thing in the run, and it produced the dominant
``heterogeneity dominates'' finding.} This is an interesting cost-of-finding
ratio: the most expensive experiment produced the most consequential result.

\subsection{What survives all $10$ rotations}
\begin{itemize}[leftmargin=*,nosep]
    \item \textbf{Heterogeneity dominates.} At $\beta\!=\!0.1$ no method
        works; at $\beta\!=\!\infty$ all methods work; the algorithm choice
        is second-order to the data choice.
    \item \textbf{\fedprox{}-$\mu\!=\!0.1$ is the natural-clients winner}
        on AUPRC, AUROC, worst-recall, and reliability simultaneously, with
        the price of $\sim\!7$ accuracy points vs.\ centralized.
    \item \textbf{The federation pays for the smallest clients.} Local-only
        on Neuro Intermediate scores AUPRC $0.11$; federated scores $0.31$;
        the $+0.20$ is what federation literally exists for.
    \item \textbf{LP shadow price is informative only inside the active region.}
        Above the cliff, more bytes do not help.
    \item \textbf{Logistic confirms MLP qualitatively}, with a $0.05$ AUPRC
        gap at $100\!\times$ less communication. The MLP is an upgrade, not
        a necessity.
\end{itemize}

\subsection{What does \emph{not} survive}
\begin{itemize}[leftmargin=*,nosep]
    \item ``CVaR helps fairness.'' The CVaR-$\alpha\!\geq\!0.75$ block is a
        discreteness collapse; CVaR-$0$ is unreliable; the empirical fairness
        gain attributed to CVaR in the original run is statistically not
        there.
    \item ``GA strictly beats grid.'' True for our grid mesh; would close to
        a tie at finer mesh.
    \item ``Centralized is the best model.'' Only on accuracy. On every
        rare-class metric and on worst-recall it loses.
\end{itemize}
