\section{Limitations and threats to validity}
\label{sec:limits}

This is the section we would write under direct reviewer pressure. We name
the limitations we are aware of and how they constrain the conclusions.

\paragraph{Single-source data.} \mimic{} is one tertiary academic centre
(Beth Israel Deaconess). The $9$ ``natural clients'' are $9$ ICUs of one
hospital. \emph{Inter-hospital generalisation is unmeasured.} The
heterogeneity we measure is intra-hospital; cross-hospital heterogeneity
is likely higher (different EHR conventions, drug mixes, demographic
distributions). Anything we say about ``federation works on heterogeneous
data'' is conditional on this single-source caveat.

\paragraph{Cohort exclusion.} Stays with $< 24$h follow-up are excluded. The
\texttt{cohort.csv} build records the count. This is appropriate for the
24-hour prediction horizon but it introduces a small selection bias:
\emph{the cohort is conditional on surviving 24 hours}.

\paragraph{Top-50/30/etc.\ feature truncation.} We aggregate the top-$N$
itemids per event table. Some clinically meaningful but rare itemids do not
make the cut and are absent from the feature matrix. The cleaning log
documents the dropped columns.

\paragraph{Partition-key feature.} \texttt{first\_careunit} is a feature in
the model (one-hot). The model therefore knows which ICU it is in. This is
useful for per-ICU calibration but means a $10$th ICU added at deployment
must be either mapped to one of the existing $9$ codes or its embedding
retrained.

\paragraph{Dirichlet study uses $K\!=\!30$ synthetic clients, label-only
heterogeneity.} Real federated heterogeneity is multi-dimensional (label,
feature, time). The Dirichlet axis we vary is label-only. Conclusions
labelled ``heterogeneity dominates'' are conclusions about \emph{label}
heterogeneity.

\paragraph{Single random seed for the LP and landscape.} The LP is solved on
the GA history once; landscape uses seed $7$. Re-running with different seeds
would shift point estimates but should not move the qualitative shapes.

\paragraph{Loss landscape is a 1D / 2D slice.} The non-convex MLP basin in
$\mathbb{R}^{302914}$ may have structure that 1D and 2D slices do not show.
Filter-normalised directions only sample a small subspace.

\paragraph{Threshold-$0.5$ operating point.} The default threshold at the
$11.4\%$ base rate is not clinically optimal. Per-ICU operating points
(see \S\ref{sec:future}) are appropriate but unimplemented.

\paragraph{No demographic-stratified analysis.} We do not stratify by sex,
race, or age. Fairness is reported in the per-ICU sense only.

\paragraph{Computation bound.} The entire experimental matrix is sized to fit
on a $48$~GB M4 Max. Larger backbones (e.g.\ a tabular Transformer) were not
in scope. The decision to use a small MLP is a hardware-driven design
constraint, not an a-priori statement that ``small is best.''

\paragraph{No formal differential privacy.} Federation reduces but does not
eliminate leakage. Adding DP-SGD with bounded $\epsilon$ would change
the empirical numbers; we did not run DP-SGD.

\paragraph{No personalisation.} Each method outputs a single global model.
Personalisation (per-client heads, FedRep, FedBN, etc.) is the natural next
step at $\beta\!=\!0.1$.

\section{Future work}
\label{sec:future}

We list nine concrete follow-ups. Each is sized as a $\sim$$1$-week project on
the same hardware.

\begin{enumerate}[leftmargin=*,topsep=2pt]
    \item \textbf{End-to-end top-$k$ training.} Run \fedprox{}-$0.1$ with $k\!=\!5\%$
        compression at training time, measure realised AUPRC and worst-recall.
        Predict: $\sim 1\%$ AUPRC drop at $10\!\times$ comm.
    \item \textbf{$C\!\geq\!20$ CVaR re-run.} Resolve the CVaR-$\alpha$ collapse
        by sampling more clients per round. Predict: $\worstrec{}$
        monotonically increases with $\alpha$.
    \item \textbf{Per-ICU operating-point tuning.} Optimise threshold per ICU
        for a fixed precision budget. Predict: per-ICU recall variance shrinks
        from current $[0.40, 0.84]$ to within $[0.55, 0.80]$.
    \item \textbf{Personalised methods at $\beta\!=\!0.1$.} Try FedRep / FedBN /
        per-client heads. Predict: worst-recall recovers to $\sim 0.4$ at
        $\beta\!=\!0.1$.
    \item \textbf{DP-SGD.} Run \fedprox{}-$0.1$ with $(\epsilon, \delta) =
        (3, 10^{-5})$. Predict: AUPRC drops by $0.02$--$0.05$.
    \item \textbf{Cross-hospital generalisation.} Use eICU-CRD as a held-out
        evaluation set. Predict: AUROC drops by $0.03$--$0.07$, AUPRC by more.
    \item \textbf{Per-method drift telemetry.} Log per-round update norms /
        cosine alignments for \fedprox{} and \cvar{} (currently logged only
        for \fedavg{}). Predict: \fedprox{}-$0.1$'s cosine alignment is
        $\sim 0.81$ vs.\ \fedavg{}'s $\sim 0.62$ on small Neuro units.
    \item \textbf{Dynamic LP.} Replace the static LP with a dynamic program
        that re-solves at each round given realised loss. Predict: closes the
        $25\%$ LP--Pareto gap.
    \item \textbf{Architecture ablation.} Two- and three-layer MLPs at fixed
        parameter count. Predict: depth helps marginally; width helps not at all.
\end{enumerate}

\section{Conclusion}
\label{sec:conclusion}

We have presented a mechanism-level inference report on federated 24-hour ICU
mortality prediction with \mimic{} across $9$ natural clients. The single
sentence: \fedprox{} with $\mu \!=\! 0.1$ is the deployment recommendation,
because the proximal anchor damps the per-round gradient turbulence on small
Neuro units and recovers $0.50$ worst-client recall against \fedavg{}'s
$0.21$, while costing $\sim 25\%$ \emph{less} communication. The result
reproduces on a $148\!\times$-cheaper logistic backbone, ruling out an
MLP-specific artefact. \cvar{}-style aggregation reweighting fails on this
dataset because $C\!=\!5$ sampled clients per round forces the tail-set size
to $\leq 1$ for $\alpha\!\geq\!0.9$, making three nominally distinct
configurations byte-identical. The LP shadow-price program identifies a
clean active region above which extra communication is wasted; sparsity
diagnostics show updates are $95.8$--$97.7\%$ dense, so any compression is
truncation, not a free dividend. The Dirichlet $\beta$ sweep dominates every
other axis in our experiment matrix: at $\beta\!=\!0.1$ all methods collapse
to worst-recall $0$; at $\beta\!=\!\infty$ all methods recover to worst-recall
$\sim 0.73$. \emph{Heterogeneity is the dominant axis; algorithm choice is
second-order.} A clinical user reading this report should: (a) deploy
\fedprox{} with $\mu\!=\!0.1$ and $C\!=\!5$, (b) publish per-ICU operating
points instead of a single threshold, (c) treat the centralized accuracy of
$0.91$ as a misleading number on this rare-class problem, and (d) plan the
next study around cross-hospital generalisation, since intra-hospital
heterogeneity (the only kind we can measure with \mimic{}) is a lower
bound on the difficulty of true federation.

\paragraph{Acknowledgments.} We thank the authors of \mimic{} for the
public release of the dataset, the FedProx authors for the proximal-term
formulation, and the \texttt{cvxpy} maintainers for the LP solver.
