\appendix
\section{Reproducibility appendix}
\label{sec:appendix}

This appendix is a single-page reproduction manifest. Everything in the report
can be rebuilt from it.

\paragraph{Software.} Python 3.12.0, PyTorch 2.5+, NumPy 1.26+, Pandas 2.2+,
SciPy 1.13+, cvxpy 1.4+. Apple Silicon M4 Max, $48$~GB unified memory.

\paragraph{Source paths.} All \texttt{*.csv} and \texttt{*.json} cited in the
body are under
\texttt{outputs/full\_mimic\_iv/} (preprocessing, EDA),
\texttt{outputs/full\_mimic\_iv\_training/} (original MLP run, $1000$-round
budget), and \texttt{outputs/full\_mimic\_iv\_proposal\_alignment/}
(\fedprox{}, logistic, Dirichlet, sparsity, LP, landscape).

\paragraph{Seeds.} The seed pool is $\{7, 11, 19, 23, 29, 31, 37, 41, 43,
47\}$, identical for every method.

\paragraph{To regenerate every figure.}
\begin{lstlisting}
.venv/bin/python reports/inference_neurips/build_plots.py
\end{lstlisting}
This reads the CSVs cited in this report and writes 21 PNGs to
\texttt{reports/inference\_neurips/figures/}. Each figure has a docstring
listing its source CSVs.

\paragraph{To regenerate every CSV.}
\begin{lstlisting}
.venv/bin/python experiments/run_full_mimic_iv.py        # original run
.venv/bin/python experiments/run_mimic_proposal_alignment.py  # proposal run
\end{lstlisting}
Both scripts write only to their respective output roots and never overwrite
the other's outputs.

\paragraph{KKT diagnostics (Reviewer 7's request).}
Verbatim from \texttt{outputs/full\_mimic\_iv\_proposal\_alignment/lp/kkt\_diagnostics.csv}:
all $24$ programs (12 dense $+$ 12 sparse) report \texttt{primal\_feasible=True},
\texttt{dual\_feasible=True}, \texttt{kkt\_status=pass}, complementary
slackness $\leq 8\!\times\!10^{-10}$, stationarity residual $\leq 5\!\times\!10^{-10}$.

\paragraph{Statistical artefacts.}
Effect sizes (Cohen's $d$) are in
\texttt{outputs/full\_mimic\_iv\_proposal\_alignment/stats/effect\_sizes.csv}.
Confidence intervals (BCa bootstrap, $1{,}000$ reps) are in
\texttt{outputs/full\_mimic\_iv\_proposal\_alignment/stats/proposal\_confidence\_intervals.csv}.
Paired-$t$ p-values are in
\texttt{outputs/full\_mimic\_iv\_proposal\_alignment/stats/proposal\_paired\_tests.csv}.

\paragraph{Code excerpts.} The single most important code change in the
proposal-alignment run is the proximal term's iteration over named
parameters (\texttt{flopt/fedprox.py}, lines $\sim$$70$--$78$):
\begin{lstlisting}
prox_term = 0.0
for n, p in model.named_parameters():
    prox_term = prox_term + (mu / 2.0) * ((p - global_state[n]) ** 2).sum()
loss = ce(logits, y) + prox_term
loss.backward()
\end{lstlisting}
Iterating over \texttt{named\_parameters()} (rather than
\texttt{state\_dict().items()}) excludes BatchNorm/dropout buffers from the
proximal term, which is necessary for correctness.

\paragraph{Numerical tables (excerpts).}
Table~\ref{tab:per-client-clinical-app} excerpts the per-client clinical
metrics from
\texttt{outputs/full\_mimic\_iv\_training/metrics/per\_client\_clinical\_metrics.csv}
for the best-validated MLP.

\begin{table}[h]
\centering
\caption{Per-client clinical metrics (best validated MLP).}
\label{tab:per-client-clinical-app}
\small
\begin{tabular}{lrrrrrr}
\toprule
ICU                  & Tests & Deaths & Acc.\ & \auprc{} & Sens. & Spec. \\
\midrule
CVICU                & 2893 & 123 & 0.924 & 0.530 & 0.683 & 0.935 \\
CCU                  & 2080 & 267 & 0.845 & 0.693 & 0.813 & 0.849 \\
MICU                 & 3985 & 589 & 0.776 & 0.693 & 0.835 & 0.766 \\
MICU/SICU            & 3166 & 461 & 0.802 & 0.611 & 0.777 & 0.806 \\
Neuro Intermediate   &  504 &  10 & 0.972 & 0.313 & 0.400 & 0.984 \\
Neuro Stepdown       &  255 &   5 & 0.957 & 0.439 & 0.600 & 0.964 \\
Neuro SICU           &  440 &  72 & 0.848 & 0.714 & 0.778 & 0.861 \\
SICU                 & 2798 & 333 & 0.837 & 0.643 & 0.739 & 0.851 \\
TSICU                & 2167 & 225 & 0.856 & 0.683 & 0.773 & 0.866 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Resource ledger.} Peak memory $9.28$~GB; min free $41.75$~GB; max
swap $0.61$~GB; $477$ samples; latest stage \texttt{loss\_landscape}; platform
\texttt{macOS-26.3.1-arm64-arm-64bit}. The watchdog (\texttt{flopt/resource\_watchdog.py})
never triggered. Source:
\texttt{outputs/full\_mimic\_iv\_proposal\_alignment/monitoring/resource\_summary.csv}.
