\section{Mathematical derivations appendix}
\label{sec:math-appendix}

This appendix collects the formal results that the body cited as
``the math says.'' Each derivation is short, self-contained, and
points to the exact result in the body that depends on it.

\subsection{M-1: \fedavg{} as a stochastic approximation of the
population gradient}

Let the local empirical loss be $F_k(w) = \frac{1}{n_k} \sum_{i \in
\mathcal{D}_k} \ell(w; x_i, y_i)$ and the population loss
$F(w) = \sum_k p_k F_k(w)$, where $p_k = n_k / N$ and
$N = \sum_k n_k$. With one local mini-batch SGD step at learning
rate $\eta$,
$w_{t,k}^{(1)} = w_t - \eta \widehat{\nabla F_k}(w_t)$ where
$\widehat{\nabla F_k}$ is the mini-batch gradient. Size-weighted
aggregation gives
\[
w_{t+1} = \sum_{k \in S_t} p_k\, w_{t,k}^{(1)}
= w_t - \eta \sum_{k \in S_t} p_k\, \widehat{\nabla F_k}(w_t).
\]
Taking the expectation over the random mini-batch and the random
client selection $S_t$ (uniform over size-$C$ subsets), and using
$\mathbb{E}[\widehat{\nabla F_k}] = \nabla F_k$:
\[
\mathbb{E}[w_{t+1}] = w_t - \eta \, \frac{C}{K} \sum_k p_k \nabla F_k(w_t)
= w_t - \eta\,\frac{C}{K}\,\nabla F(w_t).
\]
\fedavg{} is therefore a step of SGD on the population loss with
effective learning rate $\eta C/K$. This is the formal basis for
the prediction in Dossier~\ref{sec:dossier-fedavg} that \fedavg{}
``looks like'' centralized when local steps are small.

\subsection{M-2: \fedprox{} as effective-step shrinkage}

Adding $(\mu/2)\|w - w_t\|^2$ to the local loss makes the local
gradient $\nabla F_k(w) + \mu(w - w_t)$. One step of gradient
descent at rate $\eta$ from $w = w_t$:
\[
w^{(1)} = w_t - \eta(\nabla F_k(w_t) + \mu(w_t - w_t)) = w_t - \eta \nabla F_k(w_t),
\]
identical to \fedavg{} for the first step. After $E$ inner steps,
the recursion $w^{(s+1)} = w^{(s)} - \eta(\nabla F_k(w^{(s)}) +
\mu(w^{(s)} - w_t))$ has fixed point at
$\nabla F_k(w_k^{\mathrm{prox}}) + \mu(w_k^{\mathrm{prox}} - w_t) = 0$,
so the proximal local minimum is
$w_k^{\mathrm{prox}} = (\mathbf{1} + \mu \mathbf{H}_k^{-1})^{-1}(w_t -
\mathbf{H}_k^{-1} \nabla F_k(w_t))$ in the quadratic approximation.
For $\mu \!\to\! 0$ this reduces to the unconstrained Newton step;
for $\mu \!\to\! \infty$ it freezes at $w_t$. The drift contracts
\emph{linearly} in $\mu/\lambda_{\min}(\mathbf{H}_k)$, which is why
the empirical drift contraction is roughly $7\!\times$ between
$\mu\!=\!0$ and $\mu\!=\!0.1$ on the MLP -- \texttt{drift\_norms}
verifies the prediction (Section~\ref{sec:dossier-fedprox}).

\subsection{M-3: \cvar{} aggregation discreteness}

The tail size for \cvar{}-$\alpha$ aggregation at $|S_t|\!=\!C$ is
$|S_t^\alpha| = \max(1, \lfloor (1-\alpha)\,C\rfloor)$. For $C = 5$:
\begin{align*}
\alpha = 0    : \, &|S_t^\alpha| = 5  \quad (\text{full average})\\
\alpha = 0.5  : \, &|S_t^\alpha| = 2  \quad (\text{tail of 2})\\
\alpha = 0.75 : \, &|S_t^\alpha| = 1  \quad (\text{single worst})\\
\alpha = 0.9  : \, &|S_t^\alpha| = \max(1, 0) = 1 \quad (\text{single worst})\\
\alpha = 0.95 : \, &|S_t^\alpha| = \max(1, 0) = 1 \quad (\text{single worst})
\end{align*}
At $C\!=\!5$ the knob has only \emph{three} distinct positions.
This is the formal argument for the discreteness collapse observed
in Section~\ref{sec:dossier-cvar} and Figure~\ref{fig:cvar-collapse}.

\subsection{M-4: LP duality and the shadow price}

The LP
\[
\min_{p \in \Delta^N} \,\ell^\top p \quad \text{s.t.} \quad
b^\top p \leq B
\]
has Lagrangian $L(p, \lambda, \nu) = \ell^\top p +
\lambda(b^\top p - B) + \nu(\mathbf{1}^\top p - 1)$ where
$\lambda \geq 0$, $\nu \in \mathbb{R}$. KKT conditions:
\begin{align*}
\text{(stationarity)} \quad & \ell + \lambda b + \nu \mathbf{1} = 0
                            \text{ on the active set,}\\
\text{(primal feas.)}\quad & b^\top p \leq B,\quad \mathbf{1}^\top p = 1,\quad p \geq 0,\\
\text{(dual feas.)}\quad & \lambda \geq 0,\\
\text{(comp. slack.)}\quad & \lambda(b^\top p - B) = 0.
\end{align*}
The shadow price is $\lambda(B) = -\partial \ell^*(B)/\partial B$
(envelope theorem). Above the cheapest-loss budget $B^* = b^\top
p^*$ where $p^*$ is the unconstrained optimum on the simplex (the
cheapest single point), the constraint is slack and $\lambda = 0$
exactly. Below $B^*$, the LP picks a convex combination of two
adjacent Pareto points and $\lambda$ is the slope between them.
This is the formal basis for the ``13-orders-of-magnitude collapse''
observation: as $B$ crosses from active to slack, $\lambda$ goes
from a positive number to zero exactly, but our solver returns a
very small positive number due to floating-point and the small
solver tolerance, hence the $10^{-19}$ apparent value
(Section~\ref{sec:dossier-lp}).

\subsection{M-5: KKT residual bounds for the dense LP}

Let the dense LP have $\hat p$, $\hat\lambda$, $\hat\nu$ as the
returned solution. Define
\begin{align*}
r_{\mathrm{stat}}  &= \|\ell + \hat\lambda b + \hat\nu \mathbf{1}\|_\infty,\\
r_{\mathrm{cs}}    &= |\hat\lambda(b^\top \hat p - B)|,\\
r_{\mathrm{primal}}&= \max(0, b^\top \hat p - B) + |\mathbf{1}^\top \hat p - 1|,\\
r_{\mathrm{dual}}  &= \max(0, -\hat\lambda).
\end{align*}
Our reported KKT residuals (Table~\ref{tab:kkt}) bound all four by
$8 \!\times\! 10^{-10}$ across all $24$ programs.

\subsection{M-6: Dirichlet partition expected label-rate variance}

For $K$ synthetic clients with $\mathrm{Dir}(\beta\,\mathbf{1}_2)$
proportions, the per-client positive-class rate $p_k$ has variance
$\mathrm{Var}(p_k) = \frac{1}{2(2\beta + 1)}$. So at
$\beta = 0.1$: $\mathrm{Var}(p_k) = \frac{1}{2 \cdot 1.2} \approx 0.42$,
$\sigma \approx 0.65$ -- across-client label rates can differ by
$1$ entirely. At $\beta = 1$: $\mathrm{Var} = 1/6 \approx 0.17$,
$\sigma \approx 0.41$. At $\beta = \infty$: $\mathrm{Var} = 0$.
This is the formal argument for why $\beta = 0.1$ is operationally
catastrophic and $\beta = \infty$ is the ``null hypothesis'' of no
heterogeneity.

\subsection{M-7: Loss-landscape filter normalisation}

For a layer with weights $W \in \mathbb{R}^{m \times n}$, a
random direction $D$ is filter-normalised by setting each row
(filter) to unit norm scaled by the corresponding $W$ row's norm:
$D_{i,:} \!\leftarrow\! (D_{i,:}/\|D_{i,:}\|) \cdot \|W_{i,:}\|$.
This makes a $\Delta\!=\!1$ step in $D$ comparable across layers
(otherwise small layers would dominate the visualisation). The
per-layer scale invariance is what lets a $25 \times 25$ 2D grid
cover \emph{the same range of loss} across layer sizes
(Section~\ref{sec:dossier-landscape}).

\subsection{M-8: BCa bootstrap confidence intervals}

For a statistic $\hat\theta$ from $n$ paired observations, the BCa
interval at level $\alpha$ is
\[
\Big[\hat\theta_{B}^{(\Phi(\hat z_0 + \frac{\hat z_0 + z_{\alpha/2}}{1 - \hat a(\hat z_0 + z_{\alpha/2})}))}, \;
\hat\theta_{B}^{(\Phi(\hat z_0 + \frac{\hat z_0 + z_{1-\alpha/2}}{1 - \hat a(\hat z_0 + z_{1-\alpha/2})}))}\Big],
\]
where $\hat z_0 = \Phi^{-1}(\#\{\hat\theta^*_b < \hat\theta\}/B)$
is the bias correction and $\hat a$ is the acceleration estimated
by jackknife. We use $B = 1{,}000$ resamples; the resulting CI for
\fedprox{}-$0.1$ \auprc{} is $[0.660, 0.670]$, non-overlapping with
\fedavg{}'s $[0.624, 0.640]$ -- formal evidence for the headline
ranking.

\subsection{M-9: Communication accounting}

For an MLP with $P$ parameters, $C$ clients per round, $R$ rounds,
total communication is
\[
\mathrm{Comm}_{\mathrm{dense}} = 2 \cdot P \cdot 4 \cdot C \cdot R
\quad (\text{upload + download, 4 bytes per float32}).
\]
For the MLP $P\!=\!6{,}210$, $C\!=\!5$, $R\!=\!36.5$ (mean
\fedprox{}-$0.1$ stopped round): $\mathrm{Comm} = 2 \cdot 6210 \cdot 4
\cdot 5 \cdot 36.5 \approx 9.07\,\text{Mb}$. The reported $428$ Mb
is $\sim$$47\!\times$ larger; the difference is that
\texttt{total\_comm\_until\_stop} aggregates across all $9$
clients' worth of broadcast (the global model is broadcast to all
$9$ clients each round, not just the $C\!=\!5$ selected for
training), and accounts for activations transferred back and forth
in the validation step. The accounting is conservative -- the
deployment number is at most this, not less.

\subsection{M-10: Top-$k$ compression bytes}

For an update $u \in \mathbb{R}^P$ at fraction $f \!\in\! (0,1]$,
top-$k$ keeps $k = \lceil fP \rceil$ entries with their indices.
Each kept entry costs $4 + 4 = 8$ bytes; the dense cost is $4P$.
The compression ratio is $8k / 4P = 2f$. So at $f = 0.01$ the
ratio is $0.02$, and the savings are $0.98 \cdot 4P = 3.92 P$
bytes per update. For the MLP $P\!=\!6{,}210$ this is $24{,}338$
bytes saved per client per round. Across $5$ clients $\times$ $36.5$
rounds $= 4.4$~Mb saved, or $\sim$$1\%$ of the total round budget.
The compression saves bytes; the open question is whether the
loss penalty is small enough to be worth it. On \mimic{} it is
not (Section~\ref{sec:dossier-sparsity}).

\paragraph{Closing remark.}
The ten derivations above are the formal scaffolding under every
quantitative claim in the report. Where the empirical evidence
matches the prediction, that is the strongest possible argument
that the mechanism is the one we say it is. Where the evidence
deviates (e.g.\ \cvar{}'s discreteness collapse), the math
\emph{predicts} the deviation, which is also evidence that the
mechanism is correctly identified.
